\documentclass{ees}

\begin{document}

\eesTitlePage

\eesCriticalReport{
  – & – & – & This mass apears twice in the thematic catalogue: HerEy~8 is the original version (“v1”) from 1797 (\A1). HerEy~7 represents a shortened version (“v2”) which Eybler probably created in 1820 (\B1; the earliest performance date in this source is 1820-09-01). v2 differs from v1 as follows (see segno marks in the full score):
      In the \textit{Kyrie}, v1 comprises~220 bars. By contrast, bars 1–159 have been cancelled in v2, which therefore comprises only 61~bars. Here, these cancelled bars are reproduced as movement~1a. Note that there are slight differences between bar~160 of v1 and bar~1 of v2. Moreover, Eybler changed the tempo indication from “Andante” to “Andante sostenuto”.
      In the \textit{Gloria}, Eybler changed the tempo indication in bar~1 from “Allegro” to “Allegro vivace”. Moreover, v2 is characterized by several cuts: (1, \segnoOne) 65~bars following bar~35 have been replaced by a single bar (no.~36). (2, \segnoTwo) 18~bars following bar~67 have been replaced by five bars (no.~68–72). A sheet of paper comprising the latter bars has been secured with pins above the first six bars of v1. Eybler also wrote these bars onto a separate page at the end of \A1. (3, \segnoThree) 31~bars have been cancelled after bar~96. (4, \segnoFour) 60~bars following bar~165 have been cancelled. (5, \segnoFive) 34~bars following bar~226 have been replaced by one bar (no.~227). A strip of paper comprising the latter bar has been loosely inserted into the manuscript. Eybler also wrote this bar onto a separate page at the end of \A1. (6, \segnoSix) 5~bars following bar~248 have been replaced by one bar (no.~249). A strip of paper with the latter bar has been secured with pins above the first bar of v1. Eybler also wrote this bar onto a separate page at the end of \A1. Finally, in bars 127–154, a solo for two trb has been replaced by a solo for two fag.
      In the \textit{Credo}, Eybler replaced the end of the movement (289~bars starting from \segnoOne) by a drastically shortened version (22~bars). Here, these cancelled bars are reproduced as movement~3a.
      In the \textit{Benedictus}, Eybler introduced the following cuts for v2: (1, \segnoOne) 42~bars following bar~13; (2, \segnoTwo) 14~bars following bar~45; (3, \segnoThree) 16~bars following bar~49; and (4, \segnoFive) 14~bars following bar~92. Consequently, Eybler modified bars 14, 46, 50, and 93. \\
   \\
  2 & 179 & vl 2 & bar in \A1: \wholeNoteRest \\
  5 & 94  & vlc  & 1st \quarterNote\ in \A1: \crotchetRest \\
  116 = 31ff



    & 104 & fag 1    & 1st \eighthNote\ in \B1: a8 \\
    & 124 & clno 2   & in \B1 unison with clno 1 \\
  \midrule
  6 & 6   & ob 2     & bar missing in \B1 \\
    & 33  & clno 1   & bar in \B1: e″4–d″4 \\
    & 52  & trb 2    & bar in \B1: d′4.–d′8 \\
    & 54  & S        & 1st \quarterNote\ in \B1: c″4 \\
    & 57  & vl 2     & 4th \sixteenthNote\ in \B1: d″16 \\
    & 75  & S        & bar in \B1: g′4–g′4 \\
    & 87  & vl 1     & 1st \eighthNote\ in \B1: c′8 \\
    & 93  & S        & bar in \B1: e″4–\crotchetRest \\
    & 96  & vl 1     & last \eighthNote\ in \B1: a′16–a′16 \\
    & 101 & trb 2    & bar in \B1: \crotchetRest–c′4 \\
    & 101 & S        & bar in \B1: c″4–\crotchetRest \\
}

\eesToc{}

\eesScore

\end{document}
